Reproducibility of COVID-19 pre-prints on medRxiv in 2020

### 2021-02-20 outputs/paper/paper.Rmd
First draft done!!!!!!

---
title: "Reproducibility of COVID-19 pre-prints on medRxiv"
subtitle: "Markers of open code or data not identified in 81 per cent of pre-prints"
author: "Annie Collins^[University of Toronto.]"
thanks: "Paper presented at Toronto Workshop on Reproducibility, 26 Februrary 2021. Collins thanks CANSSI Ontario for financial support. Code and data are available at: https://github.com/anniecollins/reproducibility."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "We create a dataset of all the pre-prints published on medRxiv between between 28 January 2020 and 31 January 2021. We extract the text from these pre-prints and parse them looking for for keyword markers signalling the availability of the data and code underpinning the pre-print. We are unable to find markers of either open data or open code for 81 per cent of the pre-prints in our sample. Our paper demonstrates the need to have authors categorize the degree of openness of their pre-print as part of the medRxiv submissions process, and more broadly, the need to better integrate open science training into a wide range of fields."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

@@

```{r loadpackages, include=FALSE}
library(here)
library(knitr)
library(lubridate)
library(tidyverse)
```

```{r loaddata, include=FALSE}
med_open_data_results <- read_csv(here("outputs/data/med_open_data_results.csv"))
```

@@

Scientists use open repositories of papers to more quickly disseminate their research than is possible in traditional journals. These repositories, such as arxiv, bioRxiv, and medRxiv, are a critical component of science and many results build on the work published there, thus it is important that the results that are published are credible. Pre-print repositories, specifically medRxiv, have drawn unprecedented attention in the context of the 2019 novel coronavirus (COVID-19) pandemic and the changes it has thrust upon the scientific community [@else2020]. These repositories are not peer-reviewed, and, in general, anyone with appropriate academic credentials can submit a paper.

While neither peer-review nor credentials are a panacea nor a guarantee of quality, given the importance of these repositories, it is important that scientists impose on themselves various standards for their results. Following @weissgerber2021automated we examine papers about COVID-19 published to medRxiv throughout 2020. We search for markers of open science as indicators of reproducibility, specifically open data and open code.

@@

The remainder of this paper is structured as follows: first, we discuss the process of constructing our data set through retrieving pre-prints from the medRxiv repository and mining them for open data and open code markers. Then we will explore some key findings in this raw data and explore some logistic regression models that support these findings. Lastly we will discuss the implications of these findings in the broader context of reproducibility and science during the COVID-19 pandemic, as well as next steps to expand on our findings and questions raised in the research process.

# Data & Methodology

Our primary data set consists of information extracted from the medRxiv repository combined with output from running a sample of COVID-19-related pre-prints through the Open Data Detection in Publications (ODDPub) text mining algorithm [@citePDDPub], using the `oddpub` package [@citeODDPubpackage].

We constructed this data set by first creating a local copy of the medRxiv repository via the medRxiv API and then filtering for papers related to the COVID-19 pandemic through several keyword searches to create our sampling frame (N = 9,929, including only the most recent version of any given pre-print). This data includes the following variables for each pre-print in the repository: title, abstract, author(s), date posted, research field, DOI, version number, corresponding author, corresponding author's institutional affiliation, and published DOI (if the paper has since been published in a peer reviewed journal).

We then selected a random sample of these papers (N = 1,200) to check for open data and code markers using the ODDPub algorithm. This required downloading each paper as a PDF, converting the PDFs to text files, and conducting the open data and code detecting procedure to produce a results table indicating the presence of open data or open code markers in each paper (with a value of TRUE or FALSE for each marker and the relevant open data or open code statements when applicable). These PDFs were downloaded and converted to text files using the `medrxivr` package [@citemedrxivr].

Our final data set was formed by joining these two tables together via DOI to form a data set including all original, qualitative information for each pre-print alongside its open data and open code status and markers.

```{r monthly-papers-total, fig.cap = "Number of pre-prints related to COVID-19 posted to medRxiv"}
# Number of COVID-related papers posted each month in total
med_results <- read_csv(here::here("outputs/data/med_results.csv")) # All COVID medRxiv papers (9,929)

med_results %>% 
  group_by(month = floor_date(date, "month")) %>% 
  summarise(count=n()) %>% 
  ggplot(aes(month, count)) + 
  # geom_line(stat="identity") +
  geom_col() + 
  labs(title = "Pre-prints Posted per Month",
       subtitle = "medRxiv, Total = 9,929",
       x = "Month",
       y = "Number of Pre-prints") +
  theme_minimal()
```

@@

Broadly, we are unable to find markers of either open data or open code for around 81 per cent of the 1,200 sampled pre-prints. The remaining 19 per cent of pre-prints was comprised of 4 percent of the sample containing markers for open code, 8 percent of the sample with markers for open data, and 7 percent of the sample with markers for both open data and open code (Table \@ref(tab:summarycounts)).

```{r summarycounts}
# open data/open code count summary table
med_open_data_results %>% 
  count(is_open_data, is_open_code) %>%
  mutate(prop_total = n / sum(n)) %>%
  mutate(is_open_data = ifelse(is_open_data == 0, "No", "Yes"),
         is_open_code = ifelse(is_open_code == 0, "No", "Yes")) %>% 
  kable(col.names = c("Contains Open Data Markers", "Contains Open Code Markers", "Count", "Proportion of Total"),
        caption = "Counts and proportions of open data and code markers in our sample",
        booktabs = TRUE,
        digits = 2)
```

The distribution of total sampled pre-prints and sampled pre-prints with open data or code markers roughly follows that of COVID-19-related pre-prints posted in general (Figure \@ref(fig:monthly-papers-condition-stack)). The proportion of pre-prints with open data or code has fluctuated over time but shows no consistent overall increase or decrease throughout the course of the pandemic, nor in conjunction with increases or decreases in the total number of pre-prints posted to medRxiv (Figure \@ref(fig:monthly-proporiton-of-open). In our data set, only one pre-print was sampled for the month of January 2020. This pre-print contained neither open data nor open code markers, thus the 0 per cent rate of open data and code for this month should be treated as an outlier within the sample.

```{r monthly-papers-condition-stack, fig.cap = "Number of pre-prints related to COVID-19 sampled from medRxiv, distinguished by presence of open data or code markers"}
# Number of papers for each open data/code status per month
open_status <- 
  med_open_data_results %>% 
  select(title, date, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code)
open_status$condition[open_status$condition == 2] <- "Both"
open_status$condition[open_status$condition == 0] <- "Neither"
open_status$condition[open_status$condition == 1 & open_status$is_open_code == 1] <- "Open Code"
open_status$condition[open_status$condition == 1 & open_status$is_open_data == 1] <- "Open Data"

open_status %>% 
  ggplot() + 
  geom_bar(aes(x = factor(floor_date(date, "month")), 
               fill=factor(condition, levels = c("Neither", "Open Code", "Open Data", "Both")), 
               position = "stack")) +
  labs(fill = "Type of Marker") +
  scale_x_discrete(breaks=c("2020-01-01","2020-04-01","2020-07-01", "2020-10-01", "2021-01-01"), 
                   labels=c("Jan 2020","Apr 2020","Jul 2020", "Oct 2020", "Jan 2021")) +
  ylab("Number of Pre-prints") +
  xlab("Month") +
  labs(title="Sampled Pre-prints by Month", subtitle="medRxiv, Total = 1,200") +
  theme_minimal()
```

```{r proportion-of-open, fig.cap= "Proportion of pre-prints in sample with open data or open code markers by month"}
# Proportion of papers posted with open data or open code each month
colors <- c("Open Data" = "blue", "Open Code" = "green")
med_month_sum_cov <- 
  med_open_data_results %>% 
  group_by(month = floor_date(date, "month")) %>% 
  summarise(count=n(), prop_open_code = sum(is_open_code)/n(), prop_open_data = sum(is_open_data)/n()) %>% 
  mutate(num_days = days_in_month(month),
         daily_rate = count/num_days)
med_month_sum_cov %>%
  ggplot() +
  geom_line(data = med_month_sum_cov, aes(month, prop_open_data, color = "Open Data")) +
  geom_line(data = med_month_sum_cov, aes(month, prop_open_code, color = "Open Code")) +
  ggtitle("Proportion of Sampled Pre-prints with Open Data or Code Markers") +
  xlab("Month") +
  ylab("Proportion of Pre-prints") +
  labs(colour = "Type of Marker") + 
  theme_minimal()
```

## Publication
```{r open-published-summary}
published_open_status <- 
  med_open_data_results %>% 
  select(title, date, published, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code)
published_open_status$condition[published_open_status$condition == 2] <- "Both"
published_open_status$condition[published_open_status$condition == 0] <- "Neither"
published_open_status$condition[published_open_status$condition == 1 & published_open_status$is_open_code == 1] <- "Open Code"
published_open_status$condition[published_open_status$condition == 1 & published_open_status$is_open_data == 1] <- "Open Data"

published_open_status %>% 
  count(published, condition) %>% 
  pivot_wider(names_from = condition, values_from = n) %>%
  mutate(published = ifelse(published == 0, "No", "Yes"), `Proportion with Open Data and/or Code`= round((`Open Data` + `Open Code` + Both)/(`Open Data` + `Open Code` + Both + Neither), 2)) %>%
  rename(Published = published) %>%
  kable(caption = "Counts and proportions of open data markers by whether the pre-print was published",
        booktabs = TRUE)
```

```{r open-published-stack}
published_open_status$published[published_open_status$published == 0] <- "Published"
published_open_status$published[published_open_status$published == 1] <- "Unpublished"

published_open_status %>% 
  # mutate(published = ifelse(published == 0, "Unpublished", "Published")) +
  ggplot() + 
  geom_bar(aes(x = published, fill=factor(condition, levels = c("Neither", "Open Code", "Open Data", "Both")), position = "stack")) +
  labs(fill = "Type of Marker") +
  ylab("Number of Pre-prints") +
  ggtitle("Pre-prints by Publication Status") +
  theme_minimal()
```

We see that the proportion of pre-prints with open data or code markers among those that have been published is approximately 5 per cent lower than that of pre-prints which have not been published (Table \@ref(tab:open-published-summary)). There was a low number of papers that had markers of either open data or code in both those that were published and those that were not published (Figure \@ref(fig:open-published-stack)). It is important to note here that our data set imperfectly characterizes publication. In particular, it does not have the publication details for some papers that were published, and even if it were a perfect record, there is a publication lag (estimated at an average of 60 days) that may skew the results [@kwon2020].

## Type of Pre-print

```{r stack-by-type}
# Number of open data/code papers by keyword for different types
paper_type_keywords <- 
  med_open_data_results %>% select(title, abstract, date, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code)
paper_type_keywords$condition[paper_type_keywords$condition == 2] <- "Both"
paper_type_keywords$condition[paper_type_keywords$condition == 0] <- "Neither"
paper_type_keywords$condition[paper_type_keywords$condition == 1 & open_status$is_open_code == 1] <- "Open Code"
paper_type_keywords$condition[paper_type_keywords$condition == 1 & open_status$is_open_data == 1] <- "Open Data"
paper_type_keywords$`Machine Learning` <- str_detect( paper_type_keywords$title, "machine learning") | 
                                          str_detect( paper_type_keywords$abstract, "machine learning")
paper_type_keywords$Simulation <- str_detect( paper_type_keywords$title, "simulate|simulation|simulating") | 
                                    str_detect( paper_type_keywords$abstract, "simulate|simulation|simulating|simulated")
paper_type_keywords$modelling <- str_detect( paper_type_keywords$title, "modelling|model|models|modeling") | 
                                str_detect( paper_type_keywords$abstract, "modelling|model|models|modeling")
paper_type_keywords <- paper_type_keywords %>% 
  pivot_longer(cols = c("Machine Learning", "Simulation", "modelling"), names_to = "Type", values_to = "Value") %>%
  filter(Value == TRUE)
# Plot
paper_type_keywords %>% ggplot() + 
  geom_bar(aes(x = Type, fill=factor(condition, levels = c("Neither", "Open Code", "Open Data", "Both")), position = "stack")) +
  labs(fill = "Type of Marker") +
  scale_x_discrete(labels=c("Machine Learning", "modelling", "Simulation")) +
  ylab("Number of Pre-prints") +
  xlab("Month") +
  ggtitle("Sampled Pre-prints by Month") +
  theme_minimal()

```

```{r summary-by-type}
paper_type_keywords %>% 
  count(Type, condition) %>% 
  pivot_wider(names_from = condition, values_from = n, values_fill = 0) %>% 
  mutate(`Proportion with Open Data and/or Code`= round((`Open Data` + `Open Code` + Both)/(`Open Data` + `Open Code` + Both + Neither), 2)) %>%
  kable(booktabs = TRUE)
```

We performed some simple keyword searches within the titles and abstracts of our data set to examine potential differences in the prevalence of open data or code for specific types of papers. For pre-prints including the term "machine learning" (n = 21), the proportion of pre-prints with open data or code markers rose to 38 per cent, double that of the sample overall (Table \@ref(tab:summary-by-type)). Keywords relating to simulation appeared in 80 sampled pre-prints and keywords related to modelling appeared in over one third of our sampled pre-prints. The latter two categories did not contain pre-prints with open data or code markers at a rate much higher than the general sample, with 22 and 24 per cent of each type of paper containing either open data or open code markers compared to 19 per cent overall. Due to the significant role modelling has played in the context of COVID-19, it is likely that a simple keyword search is not exclusive enough for us to judge this subset as though it only contains pre-prints presenting their own, novel models.

@@

We run our analysis in `R` [@citeR]. We fit generalized linear models with open data and open code markers as binary variables (representing their presence or absence in a pre-print) as functions of the date at which a pre-print was posted.

The models below display in greater clarity the negligible role time appears to play in the proportion of pre-prints with open data or open code posted to medRxiv. Both models output negligibly small, non-significant coefficients on the date term, indicating limited influence of the date a pre-print was posted on its likelihood of containing open data or code markers.

```{r data-time-model}
glm(is_open_data ~ date, family = binomial, data = med_open_data_results) %>%
  summary()
```

```{r code-time-model}
glm(is_open_code ~ date, family = binomial, data = med_open_data_results) %>%
  summary()
```

We get similar results for publication status as a function of open data or open code presence, however we do see a significant (p < 0.05) coefficient indicating that the presence of open code markers may have a negative influence on a pre-print's likelihood of eventual publication. It is unrealistic to assume that this means pre-prints with open code are less likely to be published on the basis of their code alone. There may be common cause factors involved, such as a disinclination of journals to publish purely computational work as has been the case with pre-print repositories like bioRxiv [@kwon2020].

```{r data-published-model}
glm(published ~ as.factor(is_open_data), family = binomial, data = med_open_data_results) %>%
  summary()
```
```{r code-published-model}
glm(published ~ as.factor(is_open_code), family = binomial, data = med_open_data_results) %>%
  summary()
```

@@

Less than 20 per cent of sampled pre-prints indicated any open data or code markers. Of those that did contain markers, 4 per cent contained open code markers only, 8 percent contained open data markers only, and 7 per cent contained both open data and code markers.

@@

Transparency and reproducibility are two hallmarks of quality scientific research. Open data and open code contribute greatly to both in allowing the scientific community to more easily verify the authenticity of purported scientific discovery and its supporting evidence, which is especially important in cases where scientific research may quickly and directly impact clinical practice or policy such as during the COVID-19 pandemic. Among a myriad of other impacts on biomedical research, COVID-19 has dramatically increased in the popularity of pre-prints from both a production and consumption stand point. The number of COVID-19 pre-prints posted to medRxiv increased dramatically in the early stages of the pandemic while non-COVID-19 pre-print rates remained consistent. The same trends were apparent in abstracts accessed by medRxiv users, where COVID-19 pre-print abstracts were viewed over 15 times the rate of non-COVID-19 pre-print abstracts [@citeFraser]. For these reasons, it is more important now than ever to examine open science standards and reproducibility within pre-print repositories.

Open data is generally accepted to be beneficial to the scientific process, and to a paper's reproducibility potential, hence it is concerning that over 80 per cent of pre-prints in our sample contained no open data markers. This concern is slightly mitigated by recognition of challenges in working with biomedical data compared with data in other fields, notably privacy and ethics concerns when working with personal data [@Floca2014]. The COVID-19 pandemic has seen an uptake in open science practices globally, as evidenced by the creation of open data repositories such as the dashboard maintained by the Center for Systems Science and Engineering at Johns Hopkins University [@citeCSSE] or the wave of publishers who have removed pay walls from published COVID-19 research [@gill2020]. An important next step may compare open data and code availability in COVID-19 pre-prints with those from other topics posted to medRxiv to examine whether the pandemic has increased transparency from the previous norm.

>>>>

---
title: "Reproducibility of COVID-19 pre-prints on medRxiv in 2020"
subtitle: "Markers of open code or data not found in 81 per cent of pre-prints"
author: "Annie Collins^[University of Toronto.]"
thanks: "Paper presented at Toronto Workshop on Reproducibility, 26 February 2021. Collins thanks CANSSI Ontario for financial support. Code and data are available at: https://github.com/anniecollins/reproducibility."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "We create a dataset of all the pre-prints published on medRxiv between 28 January 2020 and 31 January 2021. We extract the text from these pre-prints and parse them looking for keyword markers signalling the availability of the data and code underpinning the pre-print. We are unable to find markers of either open data or open code for 81 per cent of the pre-prints in our sample. Our paper demonstrates the need to have authors categorize the degree of openness of their pre-print as part of the medRxiv submissions process, and more broadly, the need to better integrate open science training into a wide range of fields."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

@@

```{r loadpackages, include=FALSE}
library(here)
library(knitr)
library(lubridate)
library(tidyverse)
library(stargazer)
```

```{r loaddata, include=FALSE}
med_open_data_results <- read_csv(here("outputs/data/med_open_data_results.csv"))

med_results <- read_csv(here::here("outputs/data/med_results.csv")) # All COVID medRxiv papers (9,929)
```

@@

Scientists use open repositories of papers to more quickly disseminate their research than is possible in traditional journals or conference proceedings. These repositories, such as arXiv, bioRxiv, medRxiv, and SocArXiv, are a critical component of science and many results build on the work published there. Thus it is important that the results that are published are credible. Pre-print repositories, specifically medRxiv, have drawn unprecedented attention in the context of the 2019 novel coronavirus (COVID-19) pandemic and the changes it has imposed on the scientific community [@else2020]. These repositories are not peer-reviewed, and, in general, anyone with appropriate academic credentials can submit a paper.

While neither peer-review nor credentials are a panacea nor a guarantee of quality, given the importance of these repositories, it is important that scientists impose on themselves various standards for their results. Following @weissgerber2021automated we examine papers about COVID-19 published to medRxiv through to January 2021. We search for markers of open science as indicators of reproducibility, specifically open data and open code.

@@

The remainder of this paper is structured as follows: in Section \@ref(data-and-methodology) we discuss the process of constructing our dataset through retrieving pre-prints from the medRxiv repository and mining them for open data and open code markers. In Sections \@ref(model) and \@ref(results) we explore some key findings in this raw data and explore some logistic regression models that support these findings. Finally, in Section \@ref(discussion) we discuss the implications of these findings in the broader context of reproducibility and science during the COVID-19 pandemic, as well as next steps to expand on our findings and questions raised in the research process.

# Data and Methodology

## Overview

Our primary dataset consists of information extracted from the medRxiv repository combined with output from running a sample of COVID-19-related pre-prints through the Open Data Detection in Publications (ODDPub) text mining algorithm [@citePDDPub], using the `oddpub` package [@citeODDPubpackage].

We constructed this dataset by first creating a local copy of the medRxiv repository via the medRxiv API. We identified papers related to the COVID-19 pandemic through several keyword searches to create our sampling frame (N = 9,929, including only the most recent version of any given pre-print). This data includes the following information about each pre-print in the repository: title, abstract, author(s), date posted, research field, DOI, version number, corresponding author, corresponding author's institutional affiliation, and published DOI (if the paper has since been published in a peer-reviewed journal).

We then selected a random sample of these papers (N = 1,200) to check for open data and code markers using the ODDPub algorithm. This required downloading each paper as a PDF and then converting the PDFs to text files, using the `medrxivr` package [@citemedrxivr]. We then conducted the open data and open code detection procedure. This produced a results table indicating the presence of open data or open code markers in each paper (with a binary variable for each marker, and the relevant open data or open code statements when applicable).

Our final dataset was formed by joining these two tables together via DOI to form a dataset including all original, qualitative information for each pre-print alongside its open data and open code status and markers.

```{r monthly-papers-total, fig.cap = "Number of pre-prints related to COVID-19 posted to medRxiv", out.width="90%", fig.align = "center"}
med_results %>% 
  group_by(month = floor_date(date, "month")) %>% 
  summarise(count=n()) %>% 
  ggplot(aes(month, count)) + 
  # geom_line(stat="identity") +
  geom_col() + 
  labs(title = "Pre-prints posted per month",
       subtitle = "medRxiv, Total = 9,929",
       x = "Month",
       y = "Number of pre-prints") +
  theme_minimal()
```

## Summary statistics

@@

Broadly, we are unable to find markers of either open data or open code for around 81 per cent of the 1,200 sampled pre-prints. The remaining 19 per cent of pre-prints was comprised of 4 per cent of the sample containing markers for open code, 8 per cent of the sample with markers for open data, and 7 per cent of the sample with markers for both open data and open code (Table \@ref(tab:summarycounts)).

```{r summarycounts}
# open data/open code count summary table
med_open_data_results %>% 
  count(is_open_data, is_open_code) %>%
  mutate(prop_total = n / sum(n)) %>%
  mutate(is_open_data = ifelse(is_open_data == 0, "No", "Yes"),
         is_open_code = ifelse(is_open_code == 0, "No", "Yes")) %>% 
  kable(col.names = c("Open data markers", "Open code markers", "Count", "Proportion of total"),
        caption = "Counts and proportions of open data and code markers in our sample",
        booktabs = TRUE,
        digits = 2)
```

The distribution of total sampled pre-prints and sampled pre-prints with open data or code markers roughly follows that of COVID-19-related pre-prints posted in general (Figure \@ref(fig:monthly-papers-condition-stack)). The proportion of pre-prints with open data or code has fluctuated over time but shows no consistent overall increase or decrease throughout the course of the pandemic, nor in conjunction with increases or decreases in the total number of pre-prints posted to medRxiv (Figure \@ref(fig:monthly-proportion-of-open)). In our dataset, only one pre-print was sampled for the month of January 2020. This pre-print contained neither open data nor open code markers, thus the 0 per cent rate of open data and code for this month will be treated as an outlier, and removed modelling purposes.

```{r monthly-papers-condition-stack, fig.cap = "Number of pre-prints related to COVID-19 sampled from medRxiv, distinguished by presence of open data or code markers", out.width="90%", fig.align = "center"}
# Number of papers for each open data/code status per month
open_status <- 
  med_open_data_results %>% 
  select(title, date, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code)
open_status$condition[open_status$condition == 2] <- "Both"
open_status$condition[open_status$condition == 0] <- "Neither"
open_status$condition[open_status$condition == 1 & open_status$is_open_code == 1] <- "Open Code"
open_status$condition[open_status$condition == 1 & open_status$is_open_data == 1] <- "Open Data"

open_status %>% 
  ggplot() + 
  geom_bar(aes(x = factor(floor_date(date, "month")), 
               fill=factor(condition, levels = c("Neither", "Open Code", "Open Data", "Both")), 
               position = "stack")) +
  labs(fill = "Type of marker") +
  scale_x_discrete(breaks=c("2020-01-01","2020-04-01","2020-07-01", "2020-10-01", "2021-01-01"), 
                   labels=c("Jan 2020","Apr 2020","Jul 2020", "Oct 2020", "Jan 2021")) +
  ylab("Number of pre-prints") +
  xlab("Month") +
  labs(title="Sampled pre-prints by month", subtitle="medRxiv, Total = 1,200") +
  theme_minimal()
```

```{r monthly-proportion-of-open, fig.cap= "Proportion of pre-prints in sample with open data or open code markers by month", out.width="90%", fig.align = "center"}
# Proportion of papers posted with open data or open code each month
colors <- c("Open Data" = "blue", "Open Code" = "green")
med_month_sum_cov <- 
  med_open_data_results %>% 
  group_by(month = floor_date(date, "month")) %>% 
  summarise(count=n(), prop_open_code = sum(is_open_code)/n(), prop_open_data = sum(is_open_data)/n()) %>% 
  mutate(num_days = days_in_month(month),
         daily_rate = count/num_days)
med_month_sum_cov %>%
  ggplot() +
  geom_line(data = med_month_sum_cov, aes(month, prop_open_data, color = "Open data")) +
  geom_line(data = med_month_sum_cov, aes(month, prop_open_code, color = "Open code")) +
  ggtitle("Proportion of sampled pre-prints with open data or code markers") +
  labs(colour = "Type of marker",
       x = "Month",
       y = "Proportion of pre-prints") + 
  theme_minimal()
```

## Publication status

In Table \@ref(tab:open-published-summary) we disaggregate pre-prints by whether there is an indication that it was published. We find that the proportion of pre-prints with open data or code markers among those that have been published is approximately five per cent lower compared with pre-prints that have not been published (Table \@ref(tab:open-published-summary)).

```{r open-published-summary}
published_open_status <- 
  med_open_data_results %>% 
  select(title, date, published, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code)
published_open_status$condition[published_open_status$condition == 2] <- "Both"
published_open_status$condition[published_open_status$condition == 0] <- "Neither"
published_open_status$condition[published_open_status$condition == 1 & published_open_status$is_open_code == 1] <- "Open code"
published_open_status$condition[published_open_status$condition == 1 & published_open_status$is_open_data == 1] <- "Open data"

published_open_status %>% 
  count(published, condition) %>% 
  pivot_wider(names_from = condition, values_from = n) %>%
  mutate(published = ifelse(published == 0, "No", "Yes"), `Proportion with either/both`= round((`Open data` + `Open code` + Both)/(`Open data` + `Open code` + Both + Neither), 2)) %>%
  rename(Published = published) %>%
  kable(caption = "Counts and proportions of open data markers by whether the pre-print was published",
        booktabs = TRUE)
```

There was a low number of papers that had markers of either open data or code in both those that were published and those that were not published (Figure \@ref(fig:open-published-stack)).

```{r open-published-stack, fig.cap= "Number of pre-prints in sample with open data or open code markers by publication status", out.width="90%", fig.align = "center"}
published_open_status$published[published_open_status$published == 0] <- "Published"
published_open_status$published[published_open_status$published == 1] <- "Unpublished"

published_open_status %>% 
  # mutate(published = ifelse(published == 0, "Unpublished", "Published")) +
  ggplot() + 
  geom_bar(aes(x = published, fill=factor(condition, levels = c("Neither", "Open code", "Open data", "Both")), position = "stack")) +
  labs(fill = "Type of marker",
       y = "Number of pre-prints",
       x = "Publication status") +
  ggtitle("Pre-prints by publication status") +
  theme_minimal()
```

It is important to note that our dataset imperfectly characterizes publication. In particular, it does not have the publication details for some papers that were published. And even if it were a perfect record, there is a publication lag (estimated at an average of around 60 days) that may skew the results [@kwon2020].

## Type of pre-print

There are a variety of different types of pre-prints on medRxiv, including those focused on machine learning, modelling, and simulation. It may be that there are differences in open science practices between these. For instance, in papers related to simulation there is less likely to be issues around sharing code.

To characterize the pre-prints by type, we performed keyword searches within the titles and abstracts of our dataset to examine potential differences in the prevalence of open data or code for specific types of papers. There were 21 sampled pre-prints that included the term 'machine learning'. Of those, the proportion of pre-prints with open data or code markers rose to 38 per cent, double that of the sample overall (Table \@ref(tab:summary-by-type)). Similarly, keywords relating to simulation appeared in 80 sampled pre-prints and keywords related to modelling appeared in over one-third of our sampled pre-prints. The latter two categories did not contain pre-prints with open data or code markers at a rate much higher than the general sample, with 22 and 24 per cent of each type of paper containing either open data or open code markers compared to 19 per cent overall. However, due to the significant role that modelling has played in the context of COVID-19, it is likely that a simple keyword search does not have the resolution needed to properly characterize this subset, it is it possible that it only contains pre-prints presenting their own, novel, models.

```{r stack-by-type, fig.cap= "Number of pre-prints in sample with open data or open code markers by type", out.width="90%", fig.align = "center"}

# Number of open data/code papers by keyword for different types
paper_type_keywords <- 
  med_open_data_results %>% select(title, abstract, date, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code)
paper_type_keywords$condition[paper_type_keywords$condition == 2] <- "Both"
paper_type_keywords$condition[paper_type_keywords$condition == 0] <- "Neither"
paper_type_keywords$condition[paper_type_keywords$condition == 1 & open_status$is_open_code == 1] <- "Open code"
paper_type_keywords$condition[paper_type_keywords$condition == 1 & open_status$is_open_data == 1] <- "Open data"
paper_type_keywords$`Machine Learning` <- str_detect( paper_type_keywords$title, "machine learning") | 
                                          str_detect( paper_type_keywords$abstract, "machine learning")
paper_type_keywords$Simulation <- str_detect( paper_type_keywords$title, "simulate|simulation|simulating") | 
                                    str_detect( paper_type_keywords$abstract, "simulate|simulation|simulating|simulated")
paper_type_keywords$modelling <- str_detect( paper_type_keywords$title, "modelling|model|models|modeling") | 
                                str_detect( paper_type_keywords$abstract, "modelling|model|models|modeling")
paper_type_keywords <- paper_type_keywords %>% 
  pivot_longer(cols = c("Machine Learning", "Simulation", "modelling"), names_to = "Type", values_to = "Value") %>%
  filter(Value == TRUE)

# # Plot
# paper_type_keywords %>% ggplot() + 
#   geom_bar(aes(x = Type, fill=factor(condition, levels = c("Neither", "Open code", "Open data", "Both")), position = "stack")) +
#   labs(fill = "Type of marker") +
#   scale_x_discrete(labels=c("Machine learning", "Modelling", "Simulation")) +
#   ylab("Number of pre-prints") +
#   xlab("Month") +
#   ggtitle("Sampled pre-prints by type") +
#   theme_minimal()
```

```{r summary-by-type}
paper_type_keywords %>% 
  count(Type, condition) %>% 
  pivot_wider(names_from = condition, values_from = n, values_fill = 0) %>% 
  mutate(Type = if_else(Type == "modelling", "Modelling", Type),
         Type = if_else(Type == "Machine Learning", "Machine learning", Type),
         ) %>% 
  mutate(`Proportion with either/both`= round((`Open data` + `Open code` + Both)/(`Open data` + `Open code` + Both + Neither), 2)) %>%
  kable(caption = "Counts and proportions of open data markers by the type of pre-print",
        booktabs = TRUE)
```

@@

We are interested in understanding firstly whether open data and open code became more likely over time, and secondly whether the likelihood that at pre-print ends up published is related to provision of open data and open code. To examine these questions we estimate Equation \@ref(eq:first) by fitting two separate generalized linear models with open data and open code markers as binary variables (representing their presence or absence in a pre-print) that are explained as functions of the date at which a pre-print was posted.

\begin{equation} 
\mbox{Open data/code} = \beta_0 + \beta_1 \mbox{date} (\#eq:first)
\end{equation}

We then estimate Equation \@ref(eq:second), again by fitting separate generalized linear models, looking at whether a pre-print is published, based on whether it had open code/data.

\begin{equation} 
\mbox{Open data/code} = \beta_0 + \beta_1 \mbox{date} (\#eq:second)
\end{equation}

We estimate our models in the statistical programming language `R` [@citeR], using the function `glm()`.

Our estimates (Table \@ref(tab:datatimemodel)) indicate, in greater clarity, the negligible role time appears to play in the proportion of pre-prints with open data or open code posted to medRxiv. Both models output negligibly small, non-significant coefficients on the date term, indicating limited influence of the date a pre-print was posted on its likelihood of containing open data or code markers.

```{r, echo=FALSE, warning=FALSE, message=FALSE, results = 'asis'}
data_on_date <- glm(is_open_data ~ date, family = binomial, data = med_open_data_results)
code_on_date <- glm(is_open_code ~ date, family = binomial, data = med_open_data_results)

published_on_data <- glm(published ~ as.factor(is_open_data), family = binomial, data = med_open_data_results)
published_on_code <- glm(published ~ as.factor(is_open_code), family = binomial, data = med_open_data_results)

stargazer(data_on_date, code_on_date, published_on_data, published_on_code,
          header = FALSE,
          title = 'Estimation results',
          dep.var.labels = c('Open data', 'Open code', 'Published'),
          covariate.labels=c('Date', 'Has open data', 'Has open code', 'Constant'),
          report=('vc*s'),
          no.space = TRUE,
          digits=2,
          digits.extra=1,
          notes.append = TRUE,
          notes.align = "l",
          ci = FALSE,
          ci.level=0.95,
          label = "tab:datatimemodel"
          )
```

We get similar results for publication status as a function of open data or open code presence; however we do see a significant (p < 0.05) coefficient indicating that the presence of open code markers may have a negative influence on a pre-print's likelihood of eventual publication. It is unrealistic to assume that this means pre-prints with open code are less likely to be published on the basis of their code alone. There may be common cause factors involved, such as a disinclination of journals to publish purely computational work as has been the case with pre-print repositories like bioRxiv [@kwon2020].

@@

Less than 20 per cent of sampled pre-prints indicated any open data or code markers. Of those that did contain markers, 4 per cent contained open code markers only, 8 per cent contained open data markers only, and 7 per cent contained both open data and code markers.

@@

## On the role of transparency and reproducibility

Transparency and reproducibility are hallmarks of quality scientific research. Open data and open code contribute greatly to both in allowing the scientific community to more easily verify the authenticity of purported scientific discovery and its supporting evidence, which is especially important in cases where scientific research may quickly and directly impact clinical practice or policy such as during the COVID-19 pandemic. Among a myriad of other impacts on biomedical research, COVID-19 has dramatically increased the popularity of pre-prints from both a production and consumption stand-point. The number of COVID-19 pre-prints posted to medRxiv increased dramatically in the early stages of the pandemic while non-COVID-19 pre-print rates remained consistent. The same trends were apparent in abstracts accessed by medRxiv users, where COVID-19 pre-print abstracts were viewed over 15 times the rate of non-COVID-19 pre-print abstracts [@citeFraser]. For these reasons, it is more important now than ever to examine open science standards and reproducibility within pre-print repositories.

Open data is generally accepted to be beneficial to the scientific process, and to a paper's reproducibility potential, hence it is concerning that over 80 per cent of pre-prints in our sample contained no open data markers. This concern is slightly mitigated by recognition of challenges in working with biomedical data compared with data in other fields, notably privacy and ethics concerns when working with personal data [@Floca2014]. The COVID-19 pandemic has seen an uptake in open science practices globally, as evidenced by the creation of open data repositories such as the dashboard maintained by the Center for Systems Science and Engineering at Johns Hopkins University [@citeCSSE] or the large number of publishers who have removed paywalls from published COVID-19 research [@gill2020]. An important next step may compare open data and code availability in COVID-19 pre-prints with those from other topics posted to medRxiv to examine whether the pandemic has increased transparency from the previous norm.

Open code as an open science marker is much more context and field-dependent, as not all biomedical research papers will rely on computational methods for their analyses. However, in pre-prints where code comprises a large portion of the methodology or results, posting it openly to repositories like GitHub contributes greatly to a pre-print's potential reproducibility. This gains importance as computational methods become increasingly popular in the rush to form predictions about emerging situations with limited data or laboratory research, which was the case for modelling studies in the early days of the COVID-19 pandemic (CITE). We also see growing concern over the quality and consequences of this sort of research, with bioRxiv not allowing purely computational work [@kwon2020].

## The role of of pre-print repositories
